---
title: "Chapter 5 - Simulation"
output: html_notebook
---

# Simulation of discrete probability models

How many girls in 400 births?

```{r}
n_girls <- rbinom(n = 1, size = 400, prob = 0.488)
n_girls
```

To look at the distribution of what could happen, we simulate rhe process 1000 times:

```{r}
n_sims <- 1000
n_girls <- rep(NA, n_sims)
for (s in 1:n_sims) {
  n_girls[s] <- rbinom(n = 1, size = 400, prob = 0.488)
}
hist(n_girls)

# or
n_girls <- rbinom(n = 1000, size = 400, prob = 0.488)
hist(n_girls)
```

## Accounting for twins

```{r}
birth_type <- sample(x = c("fraternal twin", "identical twin", "single birth"),
                     size = 400, replace = TRUE, 
                     prob = c(1/125, 1/300, 1 - 1/125 - 1/300))
girls <- rep(NA, 400)
for (i in 1:400) {
  if (birth_type[i] == "single birth") {
    girls[i] <- rbinom(1, 1, 0.488)
  } else if (birth_type[i] == "identical twin") {
    girls[i] <- 2 * rbinom(1, 1, 0.495)
  } else if (birth_type[i] == "fraternal twin") {
    girls[i] <- rbinom(1, 2, 0.495)
  }
}
n_girls <- sum(girls)

# or
girls <- ifelse(birth_type == "single birth", rbinom(400, 1, 0.488),
                ifelse(birth_type == "identical twin", 2 * rbinom(400, 1, 0.495),
                       rbinom(400, 2, 0.495)))

# now simulate the distribution of the number of girls in 400 births
n_sims <- 1000
n_girls <- rep(NA, n_sims)
for (s in 1:n_sims) {
  birth_type <- sample(x = c("fraternal twin", "identical twin", "single birth"),
                     size = 400, replace = TRUE, 
                     prob = c(1/125, 1/300, 1 - 1/125 - 1/300))
  girls <- rep(NA, 400)
  for (i in 1:400) {
    if (birth_type[i] == "single birth") {
    girls[i] <- rbinom(1, 1, 0.488)
  } else if (birth_type[i] == "identical twin") {
    girls[i] <- 2 * rbinom(1, 1, 0.495)
  } else if (birth_type[i] == "fraternal twin") {
    girls[i] <- rbinom(1, 2, 0.495)
  }
  }
  n_girls[s] <- sum(girls)
}
hist(n_girls)
```

# Simulation of continuous and mixed discrete/continuous models

```{r}
n_sims <- 1000
y1 <- rnorm(n_sims, mean = 3, sd = 0.5)
y2 <- exp(y1)
y3 <- rbinom(n_sims, size = 20, prob = 0.6)
y4 <- rpois(n_sims, lambda = 5)

library(tidyverse)
p1 <- ggplot() + geom_histogram(aes(x = y1))
p2 <- ggplot() + geom_histogram(aes(x = y2)) 
p3 <- ggplot() + geom_histogram(aes(x = y3)) 
p4 <- ggplot() + geom_histogram(aes(x = y4)) 

library(patchwork)
(p1 + p2) / (p3 + p4)
```

We can also incorporate continous probability distribution with the sorts of simulations discussed in the previous section. Here is the height of one randomly chosen adult:

```{r}
male <- rbinom(1, 1, 0.48)
height <- ifelse(male == 1, rnorm(1, 69.1, 2.9), rnorm(1, 63.7, 2.7))
height
```

Now let's choose 10 adults at random and find their average height:

```{r}
N <- 10
male <- rbinom(1, 1, 0.48)
height <- ifelse(male == 1, rnorm(1, 69.1, 2.9), rnorm(1, 63.7, 2.7))
avg_height <- mean(height)
avg_height

# simulate the distribution of their heights
n_sims <- 1000
avg_height <- rep(NA, n_sims)
for (s in 1:n_sims) {
  N <- 10
  male <- rbinom(N, 1, 0.48)
  height <- ifelse(male == 1, rnorm(1, 69.1, 2.9), rnorm(1, 63.7, 2.7))
  avg_height[s] <- mean(height)
}
hist(avg_height)
```

## Simulation in R using custom-made functions

```{r}
height_sim <- function(N) {
  male <- rbinom(N, 1, 0.48)
  height <- ifelse(male == 1, rnorm(1, 69.1, 2.9), rnorm(1, 63.7, 2.7))
  mean(height)
}

# simulate
avg_height <- replicate(n = 1000, height_sim(N = 10))
hist(avg_height)
```

# Bootstrapping to simulate a sampling distribution

Bootstrap sampling is done with replacement

```{r}
earnings <- read_csv("Earnings/data/earnings.csv")
earn <- earnings$earn
male <- earnings$male
ratio <- median(earn[male == 0]) / median(earn[male == 1])
ratio
```

This shows that the median earnings of women are 60% that of men. But what about uncertianty? One approach is to look at the variation in 100 random bootstrap samples from the data, which can be taken as an approximation to the sampling distribution. 

```{r}
n <- nrow(earnings)
boot <- sample(x = n, replace = TRUE)
earn_boot <- earn[boot]
male_boot <- male[boot]
ratio_boot <- median(earn_boot[male_boot == 0]) / median(earn_boot[male_boot == 1])

boot_ratio <- function(data) {  
  n <- nrow(data)  
  boot <- sample(n, replace=TRUE)  
  earn_boot <- data$earn[boot]  
  male_boot <- data$male[boot]  
  median(earn_boot[male_boot==0])/ median(earn_boot[male_boot==1])  
}

n_sims <- 1000
output <- replicate(n_sims, boot_ratio(data = earnings))
hist(output)
```

