---
title: "Chapter 1 - Overview"
output: html_notebook
---

# Three challenges of statistics

1. generalizing from sample to population

2. generalizing from treatment to control group

3. generalizing from observed measurements to the underlying constructs of interest

# Why learn regression?

* regression is a method that allows researchers to summarize how predictions or average values of an outcome vary across individuals defined by a set of predictors

```{r}
library(tidyverse)

hibbs <- read.table("ElectionsEconomy/data/hibbs.dat", header = TRUE)

ggplot(data = hibbs, aes(x = growth, y = vote)) +
  geom_point() +
  labs(x = "Average recent growth in personal income",
       y = "Incumbent party's vote share")

library(rstanarm)
m1 <- stan_glm(vote ~ growth, data = hibbs)

# add regression line to plot
ggplot(data = hibbs, aes(x = growth, y = vote)) +
  geom_point() +
  geom_abline(intercept = coef(m1)[[1]], slope = coef(m1)[[2]]) +
  labs(x = "Average recent growth in personal income",
       y = "Incumbent party's vote share") +
  geom_label(label = "y = 46.3 + 3.0x", x = 3, y = 53)

print(m1)
```

# Exercises

1a. 

```{r}
heli <- read.table("Helicopters/data/helicopters.txt", header = TRUE)
```

1b. 

```{r}
ggplot()
```


